---
title: "nlmr_simulations"
author: "David Hughes"
date: "2023-07-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## Load needed packages ###
library(tidyverse)
library(broom)
library(patchwork)
library(mgcv)
library(metafor)
library(faux)

## Load my functions ###
f = list.files("functions/")

f
for(file in f){
  source(paste0("functions/",file))
}

```



## Simulations to evaluate problems in non-linear mr analyses

- what are the problems?

  1. stratifying on raw exposure data generates a bias so we must generate an iv-free exposure.
  2. when we stratify on the iv-free exposure the assumption that there is a constant genetic-exposure effect often appears to be invalidated.
  3. a doubly ranked methods was devised to reduce the invalidity of the constant genetic-exposure effect. However, this does not appear to remove the bias and creates new problems. 
  4. the standard error in the genetic-exposure effect is not constant across strata, in fact it tends to trend into either smaller or larger values, effectively saying that the accuracy of the genetic effect on exposure either increases or decreases with strata. This an not be. 


## Simulate some correlated data

```{r}
mydata <- rnorm_multi(n = 25000, 
                      mu = c(3, 27, 15),
                      sd = c(2, 5, 4),
                      r = c(0.3, 0.15, 0.55), 
                      varnames = c("g", "x", "y"),
                      empirical = FALSE)

cor(as.matrix(mydata))^2


```

## now add transformed versions of the the exposure normal distribution 

  - a uniform distribution
  - a gamma distribution

```{r}
## UNIFORM
a = mydata$x
a = (a - mean(a) )/ sd(a)
mydata$x_uni = pnorm( a, mean = 0, sd = 1)

## GAMMA
a = a + abs( min(a) )
mydata$x_gamma = pgamma( a , shape = 3)

## EXPONENTIAL
mydata$x_exp = pexp(mydata$x_uni, rate = 4)

## BETA
a = a / max(a)
mydata$x_beta = pbeta(a, shape1 = 4, shape2 = 3)

```



```{r, fig.width = 15, fig.height = 10}
par(mfrow = c(3,3))
hist(mydata$g, main = "PGS")
hist(mydata$x, main = "exposure")
hist(mydata$y, main = "outcome")
hist(mydata$x_uni, main = "exposure uniform")
hist(mydata$x_gamma, main = "exposure gamma")
hist(mydata$x_exp, main = "exposure exponential")
hist(mydata$x_beta, main = "exposure beta")
```



```{r}
cor(as.matrix(mydata))^2
```



# OK EVERYTHING HAS BEEN WRITTEN INTO FUNCTIONS


## NORMAL Distribution

```{r}
## simulate data
test = sim_mydata(n = 100000, seed = 546372)
## generate the summary statistics
test_ss = generate_gx_sumstats( data = test, exposure = "x", k = 20 ) ## THIS RETURNS A LIST 
        
## redefine data set
test = test_ss$wdata
## define sumstats tibble
ss = test_ss$ss

## make figures
myplot = generate_plots(ss_data = ss, simdata = test, exposure = "x", distribution_type = "normal")

## print to file
pdf(file = "figures/Simulated_Gene_Exp_Normal_Data.pdf", width = 13, height = 15)
myplot
dev.off()
```


```{r, fig.width = 13, fig.height = 15}
# Take a look at the generated figure
myplot
```



## UNIFORM Distribution

```{r}
## simulate data
test = sim_mydata(n = 100000, seed = 546372)
## generate the summary statistics
test_ss = generate_gx_sumstats( data = test, exposure = "x_uni", k = 20 )
## redefine data set
test = test_ss$wdata
## define sumstats tibble
ss = test_ss$ss
## make figures
myplot = generate_plots(ss_data = ss, simdata = test, exposure = "x_uni", distribution_type = "uniform")
## print to file
pdf(file = "figures/Simulated_Gene_Exp_Uniform_Data.pdf", width = 13, height = 15)
myplot
dev.off()
```

## EXPONENTIAL Distribution

```{r}
## simulate data
test = sim_mydata(n = 100000, seed = 546372)
## generate the summary statistics
test_ss = generate_gx_sumstats( data = test, exposure = "x_exp", k = 20 )
## redefine data set
test = test_ss$wdata
## define sumstats tibble
ss = test_ss$ss
## make figures
myplot = generate_plots(ss_data = ss, simdata = test, exposure = "x_exp", distribution_type = "exponential")
## print to file
pdf(file = "figures/Simulated_Gene_Exp_Exponential_Data.pdf", width = 13, height = 15)
myplot
dev.off()
```



## BETA Distribution

```{r}
## simulate data
test = sim_mydata(n = 100000, seed = 546372)
## generate the summary statistics
test_ss = generate_gx_sumstats( data = test, exposure = "x_beta", k = 20 )
## redefine data set
test = test_ss$wdata
## define sumstats tibble
ss = test_ss$ss
## make figures
myplot = generate_plots(ss_data = ss, simdata = test, exposure = "x_beta", distribution_type = "beta")
## print to file
pdf(file = "figures/Simulated_Gene_Exp_Beta_Data.pdf", width = 13, height = 15)
myplot
dev.off()
```

Now, let's look at outcomes:

```{r}
## UNIFORM
a = mydata$y
a = (a - mean(a) )/ sd(a)
mydata$y_uni = pnorm( a, mean = 0, sd = 1)

## GAMMA
a = a + abs( min(a) )
mydata$y_gamma = pgamma( a , shape = 3)

## EXPONENTIAL
mydata$y_exp = pexp(mydata$y_uni, rate = 4)

## BETA
a = a / max(a)
mydata$y_beta = pbeta(a, shape1 = 4, shape2 = 3)

```

Now to generate outcome data

```{r}
par(mfrow = c(3,3))
hist(mydata$g, main = "PGS")
hist(mydata$x, main = "exposure")
hist(mydata$y, main = "outcome")
hist(mydata$y_uni, main = "outcome uniform")
hist(mydata$y_gamma, main = "outcome gamma")
hist(mydata$y_exp, main = "outcome exponential")
hist(mydata$y_beta, main = "outcome beta")
```

```{r}
test = sim_mydata_outcomes(n = 100000, seed = 54632)
test_ss = generate_all_sumstats( data = test, exposure = "x_beta", outcome = "y_exp")


## redefine data set
test = test_ss$wdata
## define sumstats tibble
ss = test_ss$ss
## make figures

myplot = generate_plots_outcome(ss_data = ss, simdata = test, exposure = "x_beta", outcome = "y_exp", distribution_type = "beta")
myplot
## print to file
pdf(file = "figures/Simulated_Gene_Outcome_Beta_Data.pdf", width = 13, height = 20)
dev.off()

```

```{r}

analyses <- crossing(
  exposures = c("x", "x_exp", "x_beta", "x_uni", "x_gamma"),
outcomes = c("y", "y_exp", "y_beta", "y_uni", "y_gamma")
)
map2(.x = analyses$exposures, .y = analyses$outcomes, seed = 2023, .f = generate_outcome_plots_individual)

```

Now to get a few sets of replicates - using a linear IV
```{r}

map2(.x = analyses$exposures, .y = analyses$outcomes, seed = 2023, reps = 100, .f = generate_outcome_plots_linear_replicates)

```

Now to do the same with a U-shaped relationship between exposure - outcome

```{r}

analyses$exposures

map2(.x = analyses$exposures, .y = analyses$outcomes, seed = 2324, reps = 100, .f = generate_outcome_plots_u_replicates)
```

Now for thresholded

```{r}

map2(.x = analyses$exposures, .y = analyses$outcomes, seed = 2021, reps = 100, .f = generate_outcome_plots_thresh_replicates)

```

Randomly selected linear

```{r}

map2(.x = analyses$exposures, .y = analyses$outcomes, seed = 2023, reps = 100, .f = generate_outcome_plots_replicates_with_selection)
```

Now thresholded IV

```{r}
map2(.x = analyses$exposures, .y = analyses$outcomes, seed = 2023, reps = 100, .f = generate_outcome_plots_linear_replicates_thresh_iv)
```

Now for a real exposure:

```{r}
dat_bmi <- vroom::vroom("~/data/non-linear/pheno/analysis_data.tsv.gz", col_select = c("body_mass_index_bmi", "bmi_prs"))

p1 <- dat_bmi %>%
  mutate(scale = scale(body_mass_index_bmi)) %>%
  mutate(strata = generate_ranked_strata(exposure = body_mass_index_bmi, k = 9, instrument = bmi_prs, seed = 123)) %>%
  ggplot(aes(x = scale)) +
  geom_density() +
  facet_wrap(~strata) +
  xlim(-5,7.5) +
  theme_bw() +
  xlab("Body mass index (kg/m2)") +
  ylab("Doubly-ranked method")

p1
p2 <- dat_bmi %>%
  drop_na(bmi_prs, body_mass_index_bmi) %>%
  mutate(iv_free = lm(body_mass_index_bmi ~ bmi_prs, data = .,)$resid) %>%
  mutate(strata = generate_residualized_strata(iv_free, k = 9)) %>%
  ggplot(aes(x = scale(body_mass_index_bmi))) +
  geom_density() +
  facet_wrap(~strata) +
   xlim(-5,7.5) +
  theme_bw() +
  xlab("Body mass index (kg/m2)") +
  ylab("Residual method")




p3 <- dat_bmi %>%
  mutate(scale = scale(body_mass_index_bmi)) %>%
  mutate(strata = generate_ranked_strata(exposure = body_mass_index_bmi, k = 9, instrument = bmi_prs, seed = 123)) %>%
  ggplot(aes(x = body_mass_index_bmi, y = bmi_prs, colour = strata)) +
  geom_point(alpha = 0.01) + 
  geom_smooth(method = "lm") +
  xlim(12,50) +
  theme_bw() +
  xlab("Body mass index (kg/m2)") +
  ylab("Polygenic risk score for BMI")

p4 <-dat_bmi %>%
  drop_na(bmi_prs, body_mass_index_bmi) %>%
  mutate(iv_free = lm(body_mass_index_bmi ~ bmi_prs, data = .,)$resid) %>%
  mutate(strata = generate_residualized_strata(iv_free, k = 9)) %>%
    ggplot(aes(x = body_mass_index_bmi, y = bmi_prs, colour = strata)) +
  geom_point(alpha = 0.01) + 
  geom_smooth(method = "lm") +
  xlim(12,50) +
  theme_bw() +
  xlab("Body mass index (kg/m2)") +
  ylab("Polygenic risk score for BMI")


dat_bmi %>%
  drop_na(bmi_prs, body_mass_index_bmi) %>%
  mutate(iv_free = lm(body_mass_index_bmi ~ bmi_prs, data = .,)$resid) %>%
  mutate(strata = generate_residualized_strata(iv_free, k = 9)) %>%
    ggplot(aes(x = iv_free, y = bmi_prs, colour = strata)) +
  facet_wrap(~ strata) +
  geom_point(alpha = 0.01) + 
  geom_smooth(method = "lm", fullrange = T) +
  theme_bw() +
  xlab("IV-free BMI (kg/m2)") +
  ylab("Polygenic risk score for BMI") +
  xlim(-15,25)



dat_bmi_rn <- vroom::vroom("~/data/non-linear/pheno/analysis_data.tsv.gz", col_select = c("body_mass_index_bmi", "bmi_prs")) %>%
  drop_na(body_mass_index_bmi) %>%
  mutate(body_mass_index_bmi = RNOmni::RankNorm(body_mass_index_bmi))




p1rn <- dat_bmi_rn %>%
  mutate(strata = generate_ranked_strata(exposure = body_mass_index_bmi, k = 9, instrument = bmi_prs, seed = 123)) %>%
  ggplot(aes(x = body_mass_index_bmi)) +
  geom_density() +
  facet_wrap(~strata) +
  xlim(-5,7.5) +
  theme_bw() +
  xlab("Body mass index (kg/m2)") +
  ylab("Doubly-ranked method")


p2rn <- dat_bmi_rn %>%
  drop_na(bmi_prs, body_mass_index_bmi) %>%
  mutate(iv_free = lm(body_mass_index_bmi ~ bmi_prs, data = .,)$resid) %>%
  mutate(strata = generate_residualized_strata(iv_free, k = 9)) %>%
  ggplot(aes(x = body_mass_index_bmi)) +
  geom_density() +
  facet_wrap(~strata) +
  xlim(-5,7.5) +
  theme_bw() +
  xlab("Body mass index (kg/m2)") +
  ylab("Residual method")




p3rn <- dat_bmi_rn %>%

  mutate(strata = generate_ranked_strata(exposure = body_mass_index_bmi, k = 9, instrument = bmi_prs, seed = 123)) %>%
  ggplot(aes(x = body_mass_index_bmi, y = bmi_prs, colour = strata)) +
  geom_point(alpha = 0.01) + 
  geom_smooth(method = "lm") +
  xlim(-3.5,3.5) +
  theme_bw() +
  xlab("Body mass index (single normalised") +
  ylab("Polygenic risk score for BMI")

p3rn + p3
p4rn <-dat_bmi_rn %>%
  drop_na(bmi_prs, body_mass_index_bmi) %>%
  mutate(iv_free = lm(body_mass_index_bmi ~ bmi_prs, data = .,)$resid) %>%
  mutate(strata = generate_residualized_strata(iv_free, k = 9)) %>%
    ggplot(aes(x = body_mass_index_bmi, y = bmi_prs, colour = strata)) +
  geom_point(alpha = 0.01) + 
  geom_smooth(method = "lm") +
    xlim(-3.5,3.5) +
  theme_bw() +
  xlab("Body mass index (rank-transformed)") +
  ylab("Polygenic risk score for BMI")

p3rn
(p1 + p2) / (p3 + p4) + plot_layout(guides = "collect") + plot_annotation(tag_levels = "A") & theme(legend.position = 'bottom')
ggsave("bmi_examples_not_transformed.tiff", height = 10, width = 12, compression = "lzw") 


(p1rn + p2rn) / (p3rn + p4rn) + plot_layout(guides = "collect") + plot_annotation(tag_levels = "A") & theme(legend.position = 'bottom')
ggsave("bmi_examples_transformed.tiff", height = 10, width = 12, compression = "lzw") 



# doubly-doubly-ranlked



p1drn <- dat_bmi_rn %>%
  mutate(strata = generate_ranked_strata(exposure = body_mass_index_bmi, k = 9, instrument = bmi_prs, seed = 123)) %>%
  group_by(strata) %>%
  mutate(body_mass_index_bmi = RNOmni::RankNorm(body_mass_index_bmi)) %>%
  ggplot(aes(x = body_mass_index_bmi)) +
  geom_density() +
  facet_wrap(~strata) +
  xlim(-5,7.5) +
  theme_bw() +
  xlab("Body mass index (single-normalised)") +
  ylab("DPolygenic risk score for BMI")



p2drn <- dat_bmi_rn %>%
  drop_na(bmi_prs, body_mass_index_bmi) %>%
  mutate(iv_free = lm(body_mass_index_bmi ~ bmi_prs, data = .,)$resid) %>%
  mutate(strata = generate_residualized_strata(iv_free, k = 9)) %>%
  group_by(strata) %>%
    mutate(body_mass_index_bmi = RNOmni::RankNorm(body_mass_index_bmi)) %>%
  ggplot(aes(x = body_mass_index_bmi)) +
  geom_density() +
  facet_wrap(~strata) +
  xlim(-5,7.5) +
  theme_bw() +
  xlab("Body mass index (kg/m2)") +
  ylab("Residual method")




p3drn <- dat_bmi_rn %>%

  mutate(strata = generate_ranked_strata(exposure = body_mass_index_bmi, k = 9, instrument = bmi_prs, seed = 123)) %>%
    group_by(strata) %>%
    mutate(body_mass_index_bmi = RNOmni::RankNorm(body_mass_index_bmi) + mean(body_mass_index_bmi)) %>%
  ggplot(aes(x = body_mass_index_bmi, y = bmi_prs, colour = strata)) +
  geom_point(alpha = 0.01) + 
  geom_smooth(method = "lm") +
  xlim(-3.5,3.5) +
  theme_bw() +
  xlab("Body mass index (doubly-normalised)") +
  ylab("Polygenic risk score for BMI")


p3drn
p3 + p3rn + p3drn + plot_annotation(tag_levels = "A") + plot_layout(guides = "collect")
ggsave("bmi_single_double.tiff", height = 5, width = 17, compression = "lzw") 
p4rn <-dat_bmi_rn %>%
  drop_na(bmi_prs, body_mass_index_bmi) %>%
  mutate(iv_free = lm(body_mass_index_bmi ~ bmi_prs, data = .,)$resid) %>%
  mutate(strata = generate_residualized_strata(iv_free, k = 9)) %>%
    ggplot(aes(x = body_mass_index_bmi, y = bmi_prs, colour = strata)) +
  geom_point(alpha = 0.01) + 
  geom_smooth(method = "lm") +
    xlim(-3.5,3.5) +
  theme_bw() +
  xlab("Body mass index (rank-transformed)") +
  ylab("Polygenic risk score for BMI")




#  tables
```

TABLES


```{r}

dat_bmi <- vroom::vroom("~/data/non-linear/pheno/analysis_data.tsv.gz", col_select = c("body_mass_index_bmi", "bmi_prs"))
dat_bmi$body_mass_index_bmi = scale(dat_bmi$body_mass_index_bmi)

dr_raw <- dat_bmi %>%
  mutate(scale = scale(body_mass_index_bmi)) %>%
  mutate(strata = generate_ranked_strata(exposure = body_mass_index_bmi, k = 9, instrument = bmi_prs, seed = 123)) %>%
  group_by(strata) %>%
       nest() %>%
  mutate(res = map(data, .f = function(d) {tidy(lm(body_mass_index_bmi~ bmi_prs, data = d))})) %>%
  unnest(res) %>%
  filter(term == "bmi_prs") %>%
  arrange(strata) %>%
  select(estimate, std.error) %>%
  mutate(method = "untransformed", approach = "doubly-ranked")



res_raw <- dat_bmi %>%
  drop_na(bmi_prs, body_mass_index_bmi) %>%
  mutate(iv_free = lm(body_mass_index_bmi ~ bmi_prs, data = .,)$resid) %>%
  mutate(strata = generate_residualized_strata(iv_free, k = 9)) %>%
  group_by(strata) %>%
       nest() %>%
  mutate(res = map(data, .f = function(d) {tidy(lm(body_mass_index_bmi~ bmi_prs, data = d))})) %>%
  unnest(res) %>%
  filter(term == "bmi_prs") %>%
  arrange(strata) %>%
  select(estimate, std.error) %>%
  mutate(method = "untransformed", approach = "residual")

dr_sn <- dat_bmi_rn %>%

  mutate(strata = generate_ranked_strata(exposure = body_mass_index_bmi, k = 9, instrument = bmi_prs, seed = 123)) %>%
    group_by(strata) %>%
    nest() %>%
  mutate(res = map(data, .f = function(d) {tidy(lm(body_mass_index_bmi~ bmi_prs, data = d))})) %>%
  unnest(res) %>%
  filter(term == "bmi_prs") %>%
  arrange(strata) %>%
  select(estimate, std.error) %>%
  mutate(method = "single-normalised", approach = "doubly-ranked")

res_sn <- dat_bmi_rn %>%

 drop_na(bmi_prs, body_mass_index_bmi) %>%
  mutate(iv_free = lm(body_mass_index_bmi ~ bmi_prs, data = .,)$resid) %>%
  mutate(strata = generate_residualized_strata(iv_free, k = 9)) %>%
    group_by(strata) %>%
    nest() %>%
  mutate(res = map(data, .f = function(d) {tidy(lm(body_mass_index_bmi~ bmi_prs, data = d))})) %>%
  unnest(res) %>%
  filter(term == "bmi_prs") %>%
  arrange(strata) %>%
  select(estimate, std.error) %>%
  mutate(method = "single-normalised", approach = "residual")
  
dr_dn <- dat_bmi_rn %>%

  mutate(strata = generate_ranked_strata(exposure = body_mass_index_bmi, k = 9, instrument = bmi_prs, seed = 123)) %>%
    group_by(strata) %>%
    mutate(body_mass_index_bmi = RNOmni::RankNorm(body_mass_index_bmi) + mean(body_mass_index_bmi)) %>%
    nest() %>%
  mutate(res = map(data, .f = function(d) {tidy(lm(body_mass_index_bmi~ bmi_prs, data = d))})) %>%
  unnest(res) %>%
  filter(term == "bmi_prs") %>%
  arrange(strata) %>%
  select(estimate, std.error) %>%
  mutate(method = "doubly-normalised", approach = "doubly-ranked")


res_dn <- dat_bmi %>%
  drop_na(bmi_prs, body_mass_index_bmi) %>%

    mutate(iv_free = lm(body_mass_index_bmi ~ bmi_prs, data = .,)$resid) %>%
  mutate(strata = generate_residualized_strata(iv_free, k = 9)) %>%
    group_by(strata) %>%
    mutate(body_mass_index_bmi = RNOmni::RankNorm(body_mass_index_bmi[,1]) + mean(body_mass_index_bmi)) %>%
  group_by(strata) %>%
    nest() %>%
  mutate(res = map(data, .f = function(d) {tidy(lm(body_mass_index_bmi~ bmi_prs, data = d))})) %>%
  unnest(res) %>%
  filter(term == "bmi_prs") %>%
  arrange(strata) %>%
  select(estimate, std.error) %>%
  mutate(method = "doubly-normalised", approach = "residual")



bind_rows(
  res_sn,
  res_dn,
  res_raw,
  dr_raw,
  dr_sn,
  dr_dn
) %>%
  ungroup() %>%
  mutate(estimate = signif(estimate, 3), std.error = signif(std.error, 3)) %>%
  transmute(estimate = paste0(estimate, " (", std.error, ")"),
         approach = paste0(method, "_", approach), strata) %>%
  pivot_wider(values_from = estimate, names_from = approach) %>%
  gt::gt()



# figures 

dat_bmi_rn %>%

  mutate(strata = generate_ranked_strata(exposure = body_mass_index_bmi, k = 9, instrument = bmi_prs, seed = 123)) %>%
    group_by(strata) %>%
    mutate(body_mass_index_bmi_rnt = RNOmni::RankNorm(body_mass_index_bmi) + mean(body_mass_index_bmi)) %>%
  ungroup() %>%
  transmute(strata,
            bmi_prs_rnt = bmi_prs,
            body_mass_index_bmi_rnt = body_mass_index_bmi) %>%
  bind_cols(dat_bmi %>% drop_na()) %>%
  summarise(cor_bmi = cor(body_mass_index_bmi_rnt, body_mass_index_bmi))


```


Threshold at 0.5 vs gradually increasing

```{r}

p2 <- compare_thresh_exponential(exposure = "x", outcome = "y", reps = 100, seed = 2)

p2$data %>%
    # filter(beta_mr < quantile(beta_mr,0.975)) %>%
    # filter(beta_mr > quantile(beta_mr,0.025)) %>%
    filter(strata_method != "full") %>% 
    arrange(strata) %>%
    mutate(lower = beta_mr - 1.96 * se_mr, upper = beta_mr + 1.96 * se_mr) %>% 
    ggplot(aes(x = strata, y = beta_mr)) +
    geom_boxplot() +
    theme_bw() +
  geom_segment(aes(x = 0, xend = 25, y =-0, yend = 0), linetype = "dashed", color = "red", size = 1) +
    geom_segment(aes(x = 25, xend = 50, y =-0.5, yend = -0.5), linetype = "dashed", color = "red", size = 1) +
    facet_wrap(~strata_method, scales = "free") +
    labs( title = "MR effect estimates across strata" ) +
    ylab("Betas from replicates (interquartile ranges)") 
  

ggsave(filename= "figures/for_paper/threshold_effect.tiff",plot = p2, height = 5, width = 10)    
```

Now RINT

```{r}
map2(.x = analyses$exposures, .y = analyses$outcomes, seed = 2023, reps = 100, .f = generate_outcome_plots_linear_replicates_RINT)
```

Now try doing random simulations to see if meta-analysis works - yep in a simple linear situation it always works!!

```{r}
analyses <- crossing(
  exposures = c("x", "x_exp", "x_beta", "x_uni", "x_gamma"),
outcomes = c("y")
)
ma_fit <- map2_dfr(.x = analyses$exposures, .y = analyses$outcomes, seed = 2023, .f = run_linear_plots_meta_analysis, replicates = 30, n = 25000)
ma_fit2 <- map2_dfr(.x = analyses$exposures, .y = analyses$outcomes, seed = 2023, .f = run_linear_plots_meta_analysis, replicates = 30, n = 25000)
ma_fit <- map2_dfr(.x = "x", .y = "y", seed = 2023, .f = run_linear_plots_meta_analysis, replicates = 1000)
ma_fit %>%
  filter(outcome == "y") %>%
  ggplot(aes(x = true, y = beta_mr)) +
  geom_point() +
  facet_wrap(~ exposure ) +
  geom_abline() +
  ggtitle("Standard MR estimates are biased when the outcome is non-normal") +
  labs(subtitle = "Comparison of true effect estimates vs MR estimates across a number of exposure distributions") +
  theme_bw() +
  xlab("True beta") +
  ylab("Reported beta")


ggsave("sim_ma/true_vs_conventional_MR.tiff", compression = "lzw", height = 8, width = 8)


ma_fit2 %>%
 filter(outcome == "y") %>%
  ggplot(aes(x = mr_beta_fe, y = beta_mr)) +
  # geom_errorbar(aes(ymin = beta_mr - se_mr, ymax = beta_mr + se_mr)) +
  geom_point(aes(colour = method)) +
  facet_wrap(~ exposure, scales = "free") +
  geom_abline()  +
   ggtitle("Meta-analyses across strata tend to match summary estimates in simulations") +
  labs(subtitle = "Comparison of conventional MR vs fixed effects meta-analysis across strata") +
  xlab("IVW meta-analysis across strata") +
  ylab("Conventional MR estimate")

ggsave("sim_ma/IVW_vs_conventional_MR.tiff", compression = "lzw", height = 8, width = 8)


ma_fit2 %>%
 filter(outcome == "y") %>%
  ggplot(aes(x = mr_beta_re, y = beta_mr)) +
  geom_errorbar(aes(ymin = beta_mr - se_mr, ymax = beta_mr + se_mr)) +
  geom_point(aes(colour = method)) +
  facet_wrap(~ exposure, scales = "free") +
  geom_abline()  +
   ggtitle("Meta-analyses across strata tend to match summary estimates in simulations") +
  labs(subtitle = "Comparison of conventional MR vs random effects meta-analysis across strata") +
  xlab("RE meta-analysis across strata") +
  ylab("Conventional MR estimate")



ma_fit2 %>%
   filter(outcome == "y") %>%
  ggplot(aes(x = se_mr, y = mr_se_fe, colour = method)) +
  geom_point() +
  facet_wrap(~ exposure , scales = "free") +
    theme_bw() +
  theme(legend.position = "bottom") +
  xlab("Overall standard error") +
  ylab("Standard error from FE ma across strata") +
  ggtitle("SEs of FE meta-analysis are close to SEs of the main analysis but \ndiffer with differing distributions ") +
  labs(substile = "Note scale difference across strata")


ggsave("sim_ma/IVW_SEvs_conventional_MR_SE.tiff", compression = "lzw", height = 8, width = )


ma_fit2 %>%  
   filter(outcome == "y") %>%
   # filter(exposure == "x") %>%
  mutate(abs_se_diff = abs(mr_se_fe-se_mr)/se_mr) %>%
  ggplot(aes(x = abs_se_diff, y = beta_mr, colour = method)) +
  geom_point() +
  facet_wrap(~ exposure, scales = "free") +
  theme_bw() +
  xlab("Absolute SE difference (%)") +
  ylab("Beta coefficent") +
 scale_x_continuous(labels = scales::percent_format(scale = 100)) +
   ggtitle("SE's from the IVW across strata are well controlled when betas \nare relatively small ")


ggsave("sim_ma/IVW_SE_vs_IVW_betas.tiff", compression = "lzw", height = 10, width = 10)

ma_fit %>%
  mutate(abs_se_diff = abs(mr_se_fe-se_mr)/se_mr) %>%
  arrange(-abs_se_diff)

ma_fit %>%
   filter(outcome == "y" &  exposure == "x") %>%
  ggplot(aes(x = beta_gy, y = by_beta_fe, colour = method)) +
  geom_point() +
  facet_wrap(~ exposure + outcome) +
  geom_abline() +
    theme_bw() +
  theme(legend.position = "bottom") +
  xlab("Overall standard error") +
  ylab("Standard error from RE ma across strata")

ma_fit %>%
   filter(outcome == "y" &  exposure == "x") %>%
  ggplot(aes(x = beta_gy, y = by_beta_fe, colour = method)) +
  geom_point() +
  facet_wrap(~ exposure + outcome) +
  geom_abline() +
    theme_bw() +
  theme(legend.position = "bottom") +
  xlab("Overall standard error") +
  ylab("Standard error from RE ma across strata")


ma_fit %>%
   filter(outcome == "y" &  exposure == "x" & method == "ranked") %>%
  bind_rows(
    
    tibble(se_mr = 0.0631, mr_se_fe = 0.0256, method = "steve")
  ) %>%
  ggplot(aes(x = se_mr, y = mr_se_fe, colour = method)) +
  geom_point() +
  # facet_wrap(~ exposure + outcome) +
    theme_bw() +
  geom_abline() +
  geom_smooth(method = "lm") +
  theme(legend.position = "bottom") +
  xlab("Overall standard error") +
  ylab("Standard error from FE ma across strata") +
  scale_x_log10() +
  scale_y_log10()

ma_fit %>%
   group_by(exposure,outcome,method) %>%
  nest() %>%
  mutate(res = map(data, .f = function(df) {tidy(lm(se_mr ~ mr_se_fe, data = df))})) %>%
  unnest(res)

s <- ma_fit %>%
  filter(method == "ranked") %>%
   filter(outcome == "y" &  exposure == "x") %>%
  select(se_mr, mr_se_fe)

s %>%
  mutate(ratio = mr_se_fe/se_mr) %>%
  ggplot(aes(x = ratio)) +
  geom_histogram()

m <- lm(se_mr ~ mr_se_fe, data = s)
glance(m)
```




```{r}

```

