---
title: "nlmr_simulations"
author: "David Hughes"
date: "2023-07-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## Load needed packages ###
library(tidyverse)
library(broom)
library(patchwork)
library(mgcv)
library(metafor)
library(faux)

## Load my functions ###
f = list.files("functions/")

f
for(file in f){
  source(paste0("functions/",file))
}
```



## Simulations to evaluate problems in non-linear mr analyses

- what are the problems?

  1. stratifying on raw exposure data generates a bias so we must generate an iv-free exposure.
  2. when we stratify on the iv-free exposure the assumption that there is a constant genetic-exposure effect often appears to be invalidated.
  3. a doubly ranked methods was devised to reduce the invalidity of the constant genetic-exposure effect. However, this does not appear to remove the bias and creates new problems. 
  4. the standard error in the genetic-exposure effect is not constant across strata, in fact it tends to trend into either smaller or larger values, effectively saying that the accuracy of the genetic effect on exposure either increases or decreases with strata. This an not be. 


## Simulate some correlated data

```{r}
mydata <- rnorm_multi(n = 25000, 
                      mu = c(3, 27, 15),
                      sd = c(2, 5, 4),
                      r = c(0.3, 0.15, 0.55), 
                      varnames = c("g", "x", "y"),
                      empirical = FALSE)

cor(as.matrix(mydata))^2


```

## now add transformed versions of the the exposure normal distribution 

  - a uniform distribution
  - a gamma distribution

```{r}
## UNIFORM
a = mydata$x
a = (a - mean(a) )/ sd(a)
mydata$x_uni = pnorm( a, mean = 0, sd = 1)

## GAMMA
a = a + abs( min(a) )
mydata$x_gamma = pgamma( a , shape = 3)

## EXPONENTIAL
mydata$x_exp = pexp(mydata$x_uni, rate = 4)

## BETA
a = a / max(a)
mydata$x_beta = pbeta(a, shape1 = 4, shape2 = 3)

```



```{r, fig.width = 15, fig.height = 10}
par(mfrow = c(3,3))
hist(mydata$g, main = "PGS")
hist(mydata$x, main = "exposure")
hist(mydata$y, main = "outcome")
hist(mydata$x_uni, main = "exposure uniform")
hist(mydata$x_gamma, main = "exposure gamma")
hist(mydata$x_exp, main = "exposure exponential")
hist(mydata$x_beta, main = "exposure beta")
```



```{r}
cor(as.matrix(mydata))^2
```



# OK EVERYTHING HAS BEEN WRITTEN INTO FUNCTIONS


## NORMAL Distribution

```{r}
## simulate data
test = sim_mydata(n = 100000, seed = 546372)
## generate the summary statistics
test_ss = generate_gx_sumstats( data = test, exposure = "x", k = 20 ) ## THIS RETURNS A LIST 
        
## redefine data set
test = test_ss$wdata
## define sumstats tibble
ss = test_ss$ss

## make figures
myplot = generate_plots(ss_data = ss, simdata = test, exposure = "x", distribution_type = "normal")

## print to file
pdf(file = "figures/Simulated_Gene_Exp_Normal_Data.pdf", width = 13, height = 15)
myplot
dev.off()
```


```{r, fig.width = 13, fig.height = 15}
# Take a look at the generated figure
myplot
```



## UNIFORM Distribution

```{r}
## simulate data
test = sim_mydata(n = 100000, seed = 546372)
## generate the summary statistics
test_ss = generate_gx_sumstats( data = test, exposure = "x_uni", k = 20 )
## redefine data set
test = test_ss$wdata
## define sumstats tibble
ss = test_ss$ss
## make figures
myplot = generate_plots(ss_data = ss, simdata = test, exposure = "x_uni", distribution_type = "uniform")
## print to file
pdf(file = "figures/Simulated_Gene_Exp_Uniform_Data.pdf", width = 13, height = 15)
myplot
dev.off()
```

## EXPONENTIAL Distribution

```{r}
## simulate data
test = sim_mydata(n = 100000, seed = 546372)
## generate the summary statistics
test_ss = generate_gx_sumstats( data = test, exposure = "x_exp", k = 20 )
## redefine data set
test = test_ss$wdata
## define sumstats tibble
ss = test_ss$ss
## make figures
myplot = generate_plots(ss_data = ss, simdata = test, exposure = "x_exp", distribution_type = "exponential")
## print to file
pdf(file = "figures/Simulated_Gene_Exp_Exponential_Data.pdf", width = 13, height = 15)
myplot
dev.off()
```



## BETA Distribution

```{r}
## simulate data
test = sim_mydata(n = 100000, seed = 546372)
## generate the summary statistics
test_ss = generate_gx_sumstats( data = test, exposure = "x_beta", k = 20 )
## redefine data set
test = test_ss$wdata
## define sumstats tibble
ss = test_ss$ss
## make figures
myplot = generate_plots(ss_data = ss, simdata = test, exposure = "x_beta", distribution_type = "beta")
## print to file
pdf(file = "figures/Simulated_Gene_Exp_Beta_Data.pdf", width = 13, height = 15)
myplot
dev.off()
```

Now, let's look at outcomes:

```{r}
## UNIFORM
a = mydata$y
a = (a - mean(a) )/ sd(a)
mydata$y_uni = pnorm( a, mean = 0, sd = 1)

## GAMMA
a = a + abs( min(a) )
mydata$y_gamma = pgamma( a , shape = 3)

## EXPONENTIAL
mydata$y_exp = pexp(mydata$y_uni, rate = 4)

## BETA
a = a / max(a)
mydata$y_beta = pbeta(a, shape1 = 4, shape2 = 3)

```

Now to generate outcome data

```{r}
par(mfrow = c(3,3))
hist(mydata$g, main = "PGS")
hist(mydata$x, main = "exposure")
hist(mydata$y, main = "outcome")
hist(mydata$y_uni, main = "outcome uniform")
hist(mydata$y_gamma, main = "outcome gamma")
hist(mydata$y_exp, main = "outcome exponential")
hist(mydata$y_beta, main = "outcome beta")
```

```{r}
test = sim_mydata_outcomes(n = 100000, seed = 54632)
test_ss = generate_all_sumstats( data = test, exposure = "x_beta", outcome = "y_exp")


## redefine data set
test = test_ss$wdata
## define sumstats tibble
ss = test_ss$ss
## make figures

myplot = generate_plots_outcome(ss_data = ss, simdata = test, exposure = "x_beta", outcome = "y_exp", distribution_type = "beta")
myplot
## print to file
pdf(file = "figures/Simulated_Gene_Outcome_Beta_Data.pdf", width = 13, height = 20)
dev.off()

```

```{r}
generate_outcome_plots_individual(exposure = "x", outcome = "y_exp", seed = 20)
analyses <- crossing(
  exposures = c("x", "x_exp", "x_beta", "x_uni", "x_gamma"),
outcomes = c("y", "y_exp", "y_beta", "y_uni", "y_gamma")
)
map2(.x = analyses$exposures, .y = analyses$outcomes, seed = 2023, .f = generate_outcome_plots_individual)

```

Now to get a few sets of replicates
```{r}

map2(.x = g$exposures, .y = g$outcomes, seed = 2023, reps = 50, .f = generate_outcome_plots_linear_replicates)

```

Now to do the same with a U-shaped relationship between exposure - outcome

```{r}

analyses$exposures

map2(.x = analyses$exposures[21:25], .y = analyses$outcomes[21:25], seed = 2324, reps = 50, .f = generate_outcome_plots_u_replicates)
```

Now for thresholded

```{r}

map2(.x = analyses$exposures[21:25], .y = analyses$outcomes[21:25], seed = 2021, reps = 50, .f = generate_outcome_plots_thresh_replicates)

```

Randomly selected linear

```{r}
  dat <-  sim_mydata_outcomes(n = 25000, seed = 123) 




map2(.x = analyses$exposures, .y = analyses$outcomes, seed = 2023, reps = 50, .f = generate_outcome_plots_replicates_with_selection)
```


```{r}
 dat <-  sim_mydata_outcomes(n = 25000, seed = 122) 
sum_stats2 <- generate_all_sumstats(exposure = "x", outcome = "y", datain = dat)
sum_stats2 <- sum_stats2$ss
resid <- sum_stats %>%
  filter(strata_method == "ranked")
resid
 ##### Non-linearity tests #####

  p_res <- sum_stats %>%
  filter(strata_method != "full") %>%

    group_by(strata_method, replicate) %>%
    nest() %>%
    mutate(res = map(data, get_pvals)) %>%
    unnest(res) %>%
  ungroup()

ranked_ps <- p_res %>%
  filter(strata_method == "ranked")
resid_ps <- p_res %>%
  filter(strata_method == "residual")


```

